%----------Acknowledgments--------------------------------------------------------------
\addsec{Acknowledgments}

My deepest appreciation goes to Professor Lenz Belzner for his invaluable guidance, insightful feedback, and encouragement throughout the development of this work. His introduction to the topics of machine learning, decision-making algorithms and Python programming provided me with a strong foundation to build upon. Expertise and mentorship from his side have been crucial in shaping the direction and execution of my research.

I am also profoundly grateful to Research Assistant Chidvilas Karpenahalli Ramakrishna for his continuous support and guidance during the entire process of writing my thesis.  His clear explanations of complex concepts, assistance in accessing essential resources, and support in setting up the necessary computational environment greatly enriched my understanding and facilitated my experiments.

I would like to express my heartfelt gratitude to Technische Hochschule Ingolstadt and all the professors of the Autonomous Vehicle Engineering study program for providing me with a solid foundation in this field. The knowledge, skills, and insights gained throughout my studies have been invaluable, not only for the successful completion of my thesis but for my entire academic journey.

Throughout the study program, I had the privilege of gaining both theoretical knowledge and practical experience in the field of autonomous vehicle engineering. The curriculum, combined with innovative research opportunities, allowed me to develop critical thinking and problem-solving skills that I will carry forward into my professional career.

Thank you to everyone who has contributed to this work, directly or indirectly. Your support, knowledge, and encouragement have significantly influenced the successful completion of my thesis and my academic growth throughout the study program.
\vspace*{1.5cm}
		
Daniil Navodey\\

Ingolstadt, Germany
